---
title: "D607_template"
author: "RickRN"
date: "`r Sys.Date()`"
output: 
  openintro::lab_report: default
  html_document:
    number_sections: yes
---

# Background
## Assignment 5: Tidying and Transforming Data


<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

Following are the goal's for this assignment:  
-Transform data between wide and long formats using tidyr package  
-Change shapes of data frames using dplyr package  
-Perform data transformations to support downstream data analysis


</div> \hfill\break


```{r step_setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
```

# Acquire Data

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

In this instance, used the assignment's PDF document as the data source to create a csv dataset using Excel.  This dataset stored on GitHub.

</div> \hfill\break


```{r step_readcsv, echo=TRUE }
theURL = "https://raw.githubusercontent.com/CUNYSPS-RickRN/Team607-1/master/Wk05_airline_arrival_status.csv"

arrivals_df <- read_csv(theURL, col_names = TRUE)

arrivals_df

```

# Data Transformation 

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

Following three rules makes a dataset tidy: variables are in columns, observations are in rows, and values are in cells.

In this step, 3 steps performed to tidy data:
- remove "blank line" by selecting status = on time or delayed
- fill missing Airline data from previous row
- pivot_longer to create observations of each origination/destination pair

Before and after views of dataset of shown

</div> \hfill\break
```{r step_data_transformation, echo=TRUE}
# Data transformation into tidy data form

arrivals_pivot_df <- arrivals_df %>% 
  subset(Status == ("on time") | Status == ("delayed")) %>% 
  fill(Airline) %>% 
  pivot_longer(c("LAX", "PHX", "SD", "SFO", "SEA"), 
               names_to = "dest", values_to = "numflights")

arrivals_pivot_df

```

# Analysis by Airport

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

This analysis compares the per-city on-time performance for both airlines.

The first pair of graphs by airline shows the number of on-time and delayed
flights by airport.  Different shapes are used to distinguish between on-time
and delayed status.  AM West Airlines has  a high concentration of flights
at the Phoenix airport while Alaska Airlines has a high concentration of flights
at the Seattle airport.

The third graph shows the proportion of each airline's on-time and delayed flights
by airport.  One noticeable observation of the Phoenix airport is AM West Airlines has a
substantially higher proportion of on-time flights when compared with Alaska Airlines
and when compared with all other airports in the sample.  In the Seattle airport 
Alaska Airlines has a substantially higher proportion of on-time flights when compared
with AM West Airlines and when compared with all other airports in the sample.

</div> \hfill\break


```{r step_analysis_city, echo=TRUE}
# 
# plot by Airline dest/status num flights info 
ggplot(data = arrivals_pivot_df) + 
  geom_point(mapping = aes(x = dest, y = numflights, color= Airline, shape = Status)) + 
  facet_wrap(~ Airline)

# create new var prop_totalflights and calc prop of on-time by destination
flight_props2 <- arrivals_pivot_df %>% 
  group_by(dest) %>% 
  #  filter(Status == "on time") %>% 
  mutate(prop_totflights = numflights / sum(numflights))


# Compare per-city on-time performance for both airlines
ggplot(data = flight_props2) + 
  geom_point(mapping = aes(x = dest, y = prop_totflights, color = Airline, shape = Status))



```

# Analysis by Airline

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

AM West Airlines had a better performance of having approximately 89% of flights on-time
while Alaska Airlines had approximately 87% of its flights on-time.

</div> \hfill\break

```{r step_analysis_Airline, echo=TRUE}
# calculate numerator
flight_props_n <- flight_props2 %>% 
  group_by(Airline) %>% 
  filter(Status == "on time", Airline == "Alaska") %>% 
  summarize(numflights)

# calculate denominator
flight_props_d <- flight_props2 %>% 
  group_by(Airline) %>% 
  filter(Airline == "Alaska") %>% 
  summarize(numflights)


OT_Alaska <- sum(flight_props_n$numflights) /sum(flight_props_d$numflights)

cat("Alaska Airlines had ", OT_Alaska, " of flights on-time.\n ",
    "Number of on-time flights=",sum(flight_props_n$numflights),
    " out of ",sum(flight_props_d$numflights))

# Calculate numerator
flight_props_n <- flight_props2 %>% 
  group_by(Airline) %>% 
  filter(Status == "on time", Airline == "AM WEST") %>% 
  summarize(numflights)

# Calculate denominator
flight_props_d <- flight_props2 %>% 
  group_by(Airline) %>% 
  filter(Airline == "AM WEST") %>% 
  summarize(numflights)


OT_AMWEST <- sum(flight_props_n$numflights) /sum(flight_props_d$numflights)
cat("AM West Airlines had ", OT_AMWEST, " of flights on-time.\n ",
    "Number of on-time flights=",sum(flight_props_n$numflights),
    " out of ",sum(flight_props_d$numflights))

```



# Conclusion

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

Airlines with hub airports defined as those with high concentration of flights
appear to have a higher number of on-time flights.  Phoenix airport appears 
to be a hub for AM West Airlines while Seattle airport appears to be a hub
for Alaska Airlines.

AM West Airlines had a better performance of having approximately 89% of flights on-time
while Alaska Airlines had approximately 87% of its flights on-time.  AM West
airlines had more than twice the number of flights than Alaska airlines.


</div> \hfill\break
