---
title: "D607_P01_RickRN"
author: "RickRN"
date: "`r Sys.Date()`"
output: 
  openintro::lab_report: default
  html_document:
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Project 1:  Data Analysis - Chess Tournament

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

In this project, a text file containing results from a Chess Tournament of 64 players will be imported from Github, then manipulated in order to calculate a chess player's mean pre-chess tournament rating of  opponents, and finally create a .csv file for use in subsequent analysis.


Collaborators: Jered Ataky, Dan Rosenfeld, Magnus Skonberg, Rick Sughrue

</div> \hfill\break
```{r setup_libs, include=FALSE}
library(tidyverse)
```

# 1: Read text file

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

In this step, the file is read using read_csv into a dataframe where the first 4 rows containing title were removed along with all rows containing all dashes (--).

Patterns in the data observed:
1.    Each player's information contained in pairs of rows.
2.    "Top line" information in odd numered rows consists of: player's number, name, total points earned in tournament, and for each of the 7 rounds game status (Win, Loss, Draw, Unplayed) and points awarded.
3.    "Bottom line" information in even numbered rows consists of: player's home State/Province, USCF ID, Pre and post ratings, followed by other information.




Source file: <https://raw.githubusercontent.com/CUNYSPS-RickRN/DATA607/master/tournamentinfo.txt>.

</div> \hfill\break
```{r read_textfile, echo=FALSE}

theURL <- c("https://raw.githubusercontent.com/CUNYSPS-RickRN/DATA607/master/tournamentinfo.txt")

chess_V2_df <- data.frame(read_csv(file=theURL, col_names=FALSE, skip = 3, comment = "--"))  # file contains rows of all dashes, comment= will remove these rows

head(read.csv(file=theURL, header = FALSE)) # print "title rows"

```

# 2: Create data frames to receive extracted data

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">


The OutputCSV_df will contain elements needed to create specified .CSV file

The Opponents_df will contain a cross-table of each player's opponents round results


</div> \hfill\break

```{r create_df, echo=TRUE}

OutputCSV_df <- data.frame()  # Data elements needed for .CSV dataset

Opponents_df <- data.frame()  # Cross-table of each player's opponents round results

```

# 3: Define Patterns for Regular Expressions

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">


Visually identified data patterns, then defined Regular Expressions.

</div> \hfill\break

```{r define_patterns, echo=TRUE}
# Patterns in first row
str_pattern_playernum = "^([0-9]{1,2})"  # pattern for player number
str_pattern_playername = c("([A-Za-z]+\\s[A-Za-z]+)(\\s[A-Za-z]+){0,}") # pattern for player name
str_pattern_totalpts = c("([0-9]+\\.[0-9]+)") # pattern for Total Pts
str_pattern_chessrounds_p1 = c("(W|L|D)(\\s+)[0-9]{1,2}") # pattern for chess rounds
str_pattern_chessrounds_p2 = c("[0-9]{1,2}") # pattern for chess rounds numbers


# Patterns in second row
str_pattern_playerstate = "^([A-Za-z]{1,2})"  # pattern for Player State/Prov
str_pattern_playerprerating_p1 = "(R:)(\\s+)([0-9]{1,4})"  # pattern for player pre-rating
str_pattern_playerprerating_p2 = "[0-9]{1,4}"  # pattern for numeric portion player pre-rank

```


# 4: Using Patterns to Extract Data Fields

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">


Examined each pair of player rows using Regular Expressions to detect and extract data fields. From the 'top' (odd numbered)' rows: player number, name, total points, and game status for each round.  From the 'bottom' (even numbered)' rows: player's State/Province and Pre-game rating.

A "for loop" iterates through each player row pair and information loaded into OutputCSV_df and Opponents_df dataframes.



</div> \hfill\break

``` {r step_extractdata, echo=TRUE}

#  The following code can be used to detect and view patterns for 'debugging' when copied near adjacent code
#  str_detect(str_trim(strChessData), str_chess_pattern)  # Detect pattern
#  str_view_all(str_trim(strChessData), str_chess_pattern)  # view pattern



x = -1  # Initial index for odd numbered rows then increment by 2
y = 0   # prime index for even numbered rows then increment by 2


for (i in 1:64)
  {

# odd numbered rows contain player name
  x= x + 2

#  chess_V2_df$X1[x]

# Extract Player Number
  str_chess_pattern = str_pattern_playernum  # pattern for player number
  strChessData <- chess_V2_df$X1[x]
  t_playernum <- as.numeric(str_extract(str_trim(strChessData), str_chess_pattern))


# Extract Player Name
  str_chess_pattern = str_pattern_playername # pattern for player name
  c_playername <- str_extract(str_trim(strChessData), str_chess_pattern)


# Extract Total Points
  str_chess_pattern = str_pattern_totalpts # pattern for Total Pts
  c_totpoints <- (str_extract(str_trim(strChessData), str_chess_pattern))
# need to convert c_totpoints to numeric

# str_detect(str_trim(strChessData), str_chess_pattern)  # Detect chess rounds
# str_view_all(str_trim(strChessData), str_chess_pattern)

  
# Extract Chess Rounds Opponents
  str_chess_pattern = str_pattern_chessrounds_p1 # pattern for chess rounds
  chess_rounds <- unlist(str_extract_all(str_trim(strChessData), str_chess_pattern))

#  str_detect(str_trim(strChessData), str_chess_pattern)  # Detect chess rounds
#  str_view_all(str_trim(strChessData), str_chess_pattern)
#  str_locate_all(str_trim(strChessData), str_chess_pattern) # locate Rounds


# Extract on pulling out opponent numbers
  str_chess_pattern = str_pattern_chessrounds_p2 # pattern for chess rounds numbers
  chess_rounds <- as.numeric(unlist(str_extract_all(str_trim(chess_rounds), str_chess_pattern))) # Extract rounds

#  str_detect(str_trim(chess_rounds), str_chess_pattern)  # Detect chess rounds
#  str_view_all(str_trim(chess_rounds), str_chess_pattern)
#  str_locate_all(str_trim(chess_rounds), str_chess_pattern) # locate Rounds


## 

# even numbered rows have location / ranking info
  y = y + 2

# Extract Player Location (State/Province)
  str_chess_pattern = str_pattern_playerstate  # pattern for Player State/Prov
  strChessData <- chess_V2_df$X1[y]
  c_loc <- str_extract(str_trim(strChessData), str_chess_pattern)


# Extract Player Pre-rating
# pre-rating part 1
  str_chess_pattern = str_pattern_playerprerating_p1  # pattern for player pre-rating
  c_prerating <- str_extract(str_trim(strChessData), str_chess_pattern)



# pre-rating part 2
  str_chess_pattern = str_pattern_playerprerating_p2  # pattern for numeric portion player pre-rank
  c_prerating <- as.numeric(str_extract(str_trim(c_prerating), str_chess_pattern))

#  str_detect(str_trim(c_prerating), str_chess_pattern)  # Detect player pre-rating
#  str_view_all(str_trim(c_prerating), str_chess_pattern)



# put player pre-rating into a df 64 players and their pre-rating values

# put player opponents 64 players x  7 rounds chess, games played
  this_Opponents_df <- data.frame(t_playernum, chess_rounds, c_prerating=NA)
  Opponents_df <- bind_rows(Opponents_df, this_Opponents_df)  # add/"bind" row to df

# expected output ".csv file" from as a dataframe 64 players x 5 output values
  theCSV <- data.frame(t_playernum, c_playername, c_loc, c_totpoints, c_prerating, c_avg_opp_prerating=NA )

# "add"/bind_rows to dataframe
  OutputCSV_df <- bind_rows(OutputCSV_df, theCSV)  # "add"/bind row to df

}


```

# 5: Obtain each Player's Opponents' Pre-Chess Rating

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">


Obtain each player's opponents' pre-chess rating and load into Opponents_df for used in following step to calculate MEAN.

</div> \hfill\break

```{r step_lookup_ratings, echo=TRUE}

# Index into each Player's Opponents to obtain their Pre-Chess Rating

for (i in 1:nrow(Opponents_df))
  {
#  
  Opponents_df$c_prerating[i] <- OutputCSV_df$c_prerating[Opponents_df$chess_rounds[i]]
}


```

# 6: Calculate MEAN Pre-chess rating of Opponents

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">


For each player, obtain the opponents' pre-ratings to calculate a rounded MEAN.

</div> \hfill\break

```{r step_calcmean, echo=TRUE}

thisval_df <- data.frame()  # define a 'working' df

# get subset of each player's opponents then calc mean of opponents' pre-ratings
for (i in 1:nrow(OutputCSV_df))
  {
    thisval <- subset(Opponents_df, t_playernum == i, select = c_prerating)
    OutputCSV_df$c_avg_opp_prerating[i] <- round(mean(thisval$c_prerating))
    
}

```
# 7: Show sample information


<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

Using the first player's information for Gary Hua, the opponents' MEAN rating of 1605  calculated by using the pre-tournament opponents’ ratings of 1436, 1563, 1600, 1610, 1649, 1663, 1716, and dividing by the total number of games played (7).


</div> \hfill\break
```{r step_show_sample, echo=TRUE}
(display_opponents <- subset(Opponents_df, t_playernum == 1, select = c(chess_rounds, c_prerating)))

(display_mean <- mean(display_opponents$c_prerating))

(round(display_mean <- mean(display_opponents$c_prerating)))



```



# 8: Create .CSV dataset of Player Summary Chess Tournament information  

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">


Create a .CSV dataset consisting of player name, location, total points earned this tournament, player's pre-rating, and mean of opponents pre-rating.

As an example using the first player, the information would be: Gary Hua, ON, 6.0, 1794, 1605

</div> \hfill\break

```{r step_createCSV, echo=TRUE}
#
# Create .CSV after removing  temporary playernum vector
OutputCSV_df <- subset(OutputCSV_df,
      select = c(c_playername, c_loc, c_totpoints, c_prerating, c_avg_opp_prerating))  # omit "t"emporary col player num


write_csv(OutputCSV_df, path = "D607_Chess_Summary.csv")

head(OutputCSV_df)


```

# Summary

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

narrative


</div> \hfill\break
