---
title: "D607_A02_RickRN"
author: "RickRN"
date: "9/5/2020"
output: html_document
---

```{r setup_1, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## DATA607 Week 2 Assignment: SQL & R

Each month, the popular Rotten Tomatoes website ranks the 100 best movies on Netflix as ranked by the Tomatometer.

September 2020: See <https://editorial.rottentomatoes.com/guide/best-netflix-movies-to-watch-right-now/>

An informal survey was created selecting the 5th film from each of set of 10 films among the top 60.

THe Movie Survey is available at <https://www.surveymonkey.com/>.  Note: Not working but can be here.


```{r setup_2, echo=TRUE}
# 
library(RMySQL)
library(tidyverse)

```

## Setup MySQL database connection 

Prompt user for database username and password then setup database connection.

```{r setup_DBcon, echo=TRUE}
#prompt for input 

#vardb_schema <- readline(prompt = "Enter your DB schema: ")
#vardb_user <- readline(prompt = "Enter your DB username: ")
#vardb_password <- readline(prompt = "Enter your password: ")


vardb_user <- "DATA607"
vardb_password <- "hello"
vardb_schema <- "a02"

cat("Schema=", vardb_schema, " username=", vardb_user, " password=", vardb_password)



mydb = dbConnect(RMySQL::MySQL(), user=vardb_user,  password=vardb_password, port=3306, dbname=vardb_schema, host='localhost')

summary(mydb)

```

## Load popular movie data table 

The .csv data will be loaded into the 

```{r load_pop_movie_table, echo=TRUE}
# put code from SQL script here. 

pop_movie_df <- dbGetQuery(mydb, "select * from survey_popmovies")

dim(pop_movie_df)
str(pop_movie_df)
pop_movie_df


```

## Load movie survey data table 

The .csv data will be loaded into the and handle missing data

```{r load_movie_survey_table, echo=TRUE}
# put code from SQL script here. 

survey_responses_df <- dbGetQuery(mydb, "select * from survey_responses")

dim(survey_responses_df)
str(survey_responses_df)
survey_responses_df

```

## Plots of movie survey data 

add narrative 

```{r do_analysis, echo=TRUE}


survey_responses_df <- dbGetQuery(mydb, "select * from survey_responses WHERE survey_seen_movie = 1")

dim(survey_responses_df)
str(survey_responses_df)
survey_responses_df

#ggplot(data=survey_responses_df, aes(x=survey_likert, y=movie_id)) + geom_line()

g <- ggplot(survey_responses_df, aes(x = survey_likert, y = movie_id)) + geom_line() + xlab('Likert Rating') + ylab('Popular Movies') + facet_wrap(~movie_id) + labs(title = "Popular Movie Likert Results")
print(g)
```

## Summary

narrative
