---
title: "D607_A10_RickRN"
author: "RickRN"
date: "`r Sys.Date()`"
output: 
  openintro::lab_report: default
  html_document:
    number_sections: yes
---

```{r step_setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Sentiment Analysis

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

In this assignment, use the primary example code in Text Mining in R - Chapter 2 Sentiment Analysis (Silge & Robinson, 2020) and extend the code in the following manner:  
-Work with a different corpus of your choosing  
-Incorporate at least 1 additional sentiment lexicon.


The NRC Word-Emotion Association Lexicon dataset is used in this code (Mohammad & Turney, 2013).

Reference:
Silge, J., & Robinson, D. (2020). Text mining with R: A tidy approach. O'Reilly Media, Inc. Retrieved from https://www.tidytextmining.com/

Mohammad, S. M., & Turney, P. D. (2013). Crowdsourcing a word–emotion association lexicon. Computational Intelligence, 29(3), 436-465. Retrieved from https://onlinelibrary.wiley.com/doi/abs/10.1111/j.1467-8640.2012.00460.x





</div> \hfill\break




# A Journal of the Plague Year

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

A Journal of the Plague Year by Daniel Defoe from Gutenberg used as corpus.
NRC Word-Emotion Association Lexicon dataset is used in this code (Mohammad & Turney, 2013).

The corpus will be tokenized and stop words will be removed.

The NRC lexicon used to identify tokens related to "fear" and "joy" found in the corpus in a plague.

A count of words related to fear and joy are ranked by frequency.

</div> \hfill\break
```{r step_y, echo=TRUE}
# comments
library(gutenbergr)
library(tidytext)



defoe <- gutenberg_download(c(376))  # A Journal of the Plague Year by Daniel Defoe

tidy_defoe <- defoe %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words)

nrc_sentiments <- get_sentiments("nrc")

nrc_joy <- get_sentiments("nrc") %>%
  filter(sentiment == "joy")

nrc_fear <- get_sentiments("nrc") %>%
  filter(sentiment == "fear")


(defoe_joy <- tidy_defoe %>%
  inner_join(nrc_joy) %>%
  count(word, sort = TRUE))

(defoe_fear <- tidy_defoe %>%
  inner_join(nrc_fear) %>%
  count(word, sort = TRUE))



```


# Summary

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

Sentiment analysis provides a way to understand the attitudes and opinions expressed in texts.  For example, in A Journal of the Plague Year, a search was made for the sentiments of fear and joy in a plague.  Clearly, "plague" and "infection" are fear related words. However, some words can have dual meanings where "God" elicits both job and fear.  

</div> \hfill\break
