---
title: "D607_A03"
author: "RickRN"
date: "9/10/2020"
output:
  html_document:
  number_sections: yes
  toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
#library(stringr)
```

# Week 3 Assignment in Character Manipulation 

_This week's assignment comprises the following:_

1. provide code that identifies College Majors containing "DATA" or "STATISTICS"

2. provide code that transforms a set of data

3. Describe, in words, what specified expressions will match

4. Construct regular expressions to match words


```{r task1_a,echo=FALSE}
# get fivethirtyeight data


theURL <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv"

five38_df <- read.csv(file=theURL, header=TRUE, sep=",")

#dim(five38_df)

#head(five38_df)

#tail(five38_df)

#names (five38_df)

```

## Task 1: Provide code that identifies College Majors containing "DATA" or "STATISTICS"

Following is the code to identify college majors containing "DATA" or "STATISTICS".

```{r task1_b, echo=TRUE}
# find majors containing specific strings

str_to_title(str_subset(str_to_lower(five38_df$Major), str_to_lower("DATA|SCIENCES")))


```
## Task 2: Provide code that transforms a set of data

Transform this vector of data from one form into another form:

```{r task2a, echo=TRUE}
# transform set of set

(task_var <-c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry"))

(new_var <- glue::glue('c("{fruitlist}")',
  fruitlist = glue::glue_collapse(task_var, sep = '", "')))

```
## Task 3: Describe, in words, what specified expressions will match

1.	(.)\1\1
2.	"(.)(.)\\2\\1"
3.	(..)\1
4.	"(.).\\1.\\1"
5.	"(.)(.)(.).*\\3\\2\\1"

```{r task3, echo=TRUE}
# patterns

# (.)\1\1 
#	"(.)(.)\\2\\1" 
#	(..)\1
#	"(.).\\1.\\1"
#	"(.)(.)(.).*\\3\\2\\1"


# 1. I was expecting the pattern below to find 1 character group, number it as \1 then repeat it twice.  expected result aaa,bbb but didn't work
str_view(c("aaa","abc","bbb"),"(.)\1\1") # from book R4DS, maybe wrong?
# Try pattern again but using 2 backslashes, book typo?
str_view(c("aaa","abc","bbb"),"(.)\\1\\1") # try pattern with 2 backslashes


# code from slack post. not sure what \1\1 really is. ESC and something?
test <- list("777", "data\1\1", "anna", "2002", '"elle"')
str_view(test, "(.)\1\1", match = TRUE)


# 2. the below example defines 2 character groups, numbers them 1 and 2 respectively, then reverses the character groups
str_view(c("abba","baba","baab","saab"),"(.)(.)\\2\\1")

# 3. I was expecting the pattern below to find 1 character group, number it as \1 then repeat it twice.  expected result aaa,bbb but didn't work
str_view(c("aaaa","abab","bbbc"),"(..)\1")
# Try pattern again but using 2 backslashes, book typo?
str_view(c("aaaa","abab","bbbc"),"(..)\\1") # try again with 2 backslashes

# 4. The pattern below finds a character, becomes group 1, followed by any char, group 1 character, followed by any character, then followed by same character as group 1.  Expected result: abaca
str_view(c("abaca","bcdbb"),"(.).\\1.\\1")  #abaca


# 5. The pattern below  finds 3 singular character groups, numbering each group respectively as 1,2,3, followed by zero or more characters then ending with the first 3 characters in reverse order.  Expected result: abccba
str_view(c("abccba","abcdcba","abc123"),"(.)(.)(.).*\\3\\2\\1")  #abccba
```
## Task 4: Construct regular expressions to match words

1.	Start and end with the same character.
2.	Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
3.	Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

```{r task4, echo=TRUE}
#
str_subset(five38_df$Major, "^(.)((.*\\1$)|\\1$)") # start & end same char

str_subset  ("church", "([A-Za-z][A-Za-z]).*\\1")  # repeated pair of letters

str_subset("eleven", "([A-Za-z]).*\\1.*\\1") # one letter repeated in 3 places

```
## Summary

Regular expressions are useful because strings usually contain unstructured or semi-structured data, and regexps are a concise language for describing patterns in strings.